\vspace{1.0cm} \LARGE \bf Revealing the competitive dynamics of influenza viruses through evolutionary cartography \\ or \\ Charting antigenic drift and evolutionary success of seasonal influenza lineages

### 2013-01-24 flux.tex
Make Table 1 more approachable per Marc's suggestion.

We test model performance by constructing training datasets representing 90\% of the HI measurements for each of the four influenza lineages and test datasets representing the remaining 10\% of the measurements for each lineage. 
By fitting the BMDS model to the training dataset, we are able to predict HI titers in the test dataset and compare these predicted titers to observed titers. 
We begin with a BMDS analog of the model used by Smith et al.\ \cite{Smith04}, in which viruses and sera are represented as 2D locations, expected titer is relative to the maximum titer of a particular ferret serum and virus and serum locations follow a diffuse prior.
We find that this model performs well, yielding an average absolute predictive error of between 0.79 and 0.95 log$_2$ HI titers across influenza lineages (model 1; Table \ref{errortable}).
We extend this model to include a time-informed prior on virus and serum locations that assumes antigenic location along dimension 1 increases in proportion to date of virus or serum sampling, resulting in a small improvement in predictive error (model 2; Table \ref{errortable}).

%%% errortable %%%
\begin{table}[h]
	\centering
	\caption{\textbf{Average absolute prediction error of log$_2$ HI titer for test data across models and datasets.}}
	\label{errortable}	
	\makebox[\textwidth][c]{
	\begin{tabular}{ c c c c c c c c } 
	\hline
	\multicolumn{4}{c}{} 	&	\multicolumn{4}{c}{Test error}  \\
	Model	&	Dimen 	& 	Serum/virus effects 	& Location Prior					& 	A/H3N2	& 	A/H1N1	& 	B/Vic	& 	B/Yam	\\
	\hline		
	1		&	2D 		&	Fixed $\ses$ 			& Diffuse							&	0.93 	&	0.78 	&	0.80	&	0.93	\\
	2		&	2D 		&	Fixed $\ses$			& $\beta,\virussd,\serumsd$			&	0.91 	&	0.74 	&	0.77	&	0.83	\\	
	3		&	1D 		&	Fixed $\ses$			& $\beta,\virussd,\serumsd$			&	1.07	&	0.92	&	0.94	&	1.08	\\		
	4		&	3D 		&	Fixed $\ses$			& $\beta,\virussd,\serumsd$			&	0.91	&	0.74 	&	0.79	&	0.84	\\
	5		&	4D 		&	Fixed $\ses$			& $\beta,\virussd,\serumsd$			&	0.95	&	0.75 	&	0.81	&	0.86	\\
	6		&	5D 		&	Fixed $\ses$			& $\beta,\virussd,\serumsd$			&	0.99	&	0.76 	&	0.84	&	0.88	\\				
	7		&	2D 		&	Estimated $\ses, \ves$	& $\beta,\virussd,\serumsd$			&	0.77	&	0.71 	&	0.65	&	0.73	\\	
	8		&	2D 		&	Estimated $\ses, \ves$	& $\beta,\virussd,\serumsd,\tree$	&	0.76	&	0.71 	&	0.64	&	0.72	\\		
	\hline
	\end{tabular}	
	}
\end{table}

>>>>

We test model performance by constructing training datasets representing 90\% of the HI measurements for each of the four influenza lineages and test datasets representing the remaining 10\% of the measurements for each lineage. 
By fitting the BMDS model to the training dataset, we are able to predict HI titers in the test dataset and compare these predicted titers to observed titers. 
We begin with a BMDS analog of the model used by Smith et al.\ \cite{Smith04}, in which viruses and sera are represented as 2D locations, expected titer is relative to the maximum titer of a particular ferret serum and virus and serum locations follow a diffuse prior.
We find that this model performs well, yielding an average absolute predictive error of between 0.79 and 0.95 log$_2$ HI titers across influenza lineages (model 1; Table \ref{errortable}).
We extend this model to include a time-informed prior that models the process of antigenic drift, in which virus and serum locations  along dimension 1 increase with date of virus or serum sampling, resulting in a small improvement in predictive error (model 2; Table \ref{errortable}).

%%% errortable %%%
\begin{table}[h]
	\centering
	\caption{\textbf{Average absolute prediction error of log$_2$ HI titer for test data across models and datasets.}}
	\label{errortable}	
	\makebox[\textwidth][c]{
	\begin{tabular}{ c c c c c c c c c } 
	\hline
	\multicolumn{5}{c}{} 	&	\multicolumn{4}{c}{Test error}  \\
	Model	&	Dimen 	& 	Serum effects 	&	Virus effects	& Location Prior	& 	A/H3N2	& 	A/H1N1	& 	B/Vic	& 	B/Yam	\\
	\hline		
	1		&	2D 		&	Fixed 			&	None			& Uninformed		&	0.93 	&	0.78 	&	0.80	&	0.93	\\
	2		&	2D 		&	Fixed			&	None			& Drift				&	0.91 	&	0.74 	&	0.77	&	0.83	\\	
	3		&	1D 		&	Fixed			&	None			& Drift				&	1.07	&	0.92	&	0.94	&	1.08	\\		
	4		&	3D 		&	Fixed			&	None			& Drift				&	0.91	&	0.74 	&	0.79	&	0.84	\\
	5		&	4D 		&	Fixed			&	None			& Drift				&	0.95	&	0.75 	&	0.81	&	0.86	\\
	6		&	5D 		&	Fixed			&	None			& Drift				&	0.99	&	0.76 	&	0.84	&	0.88	\\				
	7		&	2D 		&	Estimated		&	Estimated		& Drift				&	0.77	&	0.71 	&	0.65	&	0.73	\\	
	8		&	2D 		&	Estimated		&	Estimated		& Diffusion/Drift	&	0.76	&	0.71 	&	0.64	&	0.72	\\		
	\hline
	\end{tabular}	
	}
\end{table}
